### 2019-08-22 ThesisJo/2_background.tex
2019-08-21

Waves are made of endless tiny energy transactions, a dance of particles effecting each other in search of equilibrium. They are caused by disturbances in the forces of nature. The drop of a rock into calm water or rapid displacement of still air cause mechanical waves, while synchronized oscillations of electric and magnetic fields cause electromagnetic waves. Regardless of whether the waves are mechanical or electromagnetic, move perpendicular or parallel to the axis of propagation, what medium they propagate through or the means by which they do so, they are---perhaps surprisingly---all governed by a partial differential equation known as the \emph{wave equation}:
%
\begin{align}
\Delta f = 
\frac{\delta^2 f}{\delta x^2} + \frac{\delta^2 f}{\delta y^2} + \frac{\delta^2 f}{\delta z^2} = \frac{1}{c}\frac{\delta^2 f}{\delta t^2},
\end{align}
%
where $\Delta$ is the Laplace operator, $f$ is field strength and $c\in\mathbb{R}^+$ is wave propagation speed. If $f$ represents the displacement of a vibrating string, the equation states that the force pulling the string towards equilibrium (right side) is proportional to the string's curvature (left side). In acoustics $f$ represents sound pressure, but since the mathematics are all the same we will simply think of $f$ as a signal: a unit-less scalar function with wave characteristics.

A few solutions to the wave equation is of particular interest, that of a plane and spherical wave
%
\begin{align}
f_\textrm{plane}(\dvec{p}^A,t)  &= A e^{j(\omega t - \vec{k} \cdot \vec{p}_{SO})} \\
f_\textrm{spheric}(\dvec{p}^A,t) &= \frac{A}{\lVert \dvec{p}_{SO}^S \rVert} e^{j(\omega t - \dvec{k}^T \lVert \dvec{p}_{SO}^A \rVert)}
\end{align}
%

One is that of plane wave and the other a  A signal source with the position $\udot{S}$ . We can represent the field observed at $\udot{A}$ as $f(\dvec{p}^A,t)$, where position is represented by $\dvec{p}^A=[x_A,y_A,z_A]^T\in\mathbb{R}^3$ and time by $t\in\mathbb{R}$,

An introduction to such notation can be found in \chapref{IV_methods}{IV}.

reference frame $\ubar{A} \triangleq (\udot{A},\uvec{A})$ to it, where $\udot{A}\in\mathbb{A}$ is its absolute position and $\uvec{A}\triangleq\big(\hat{x}_{A},\hat{y}_{A},\hat{z}_{A}\big)\in\mathbb{V}^3$ its absolute orientation (where $\hat{\cdot}$ denotes unit length)

@@

Imagine we record a wavefield with $M$ sensors arranged in a uniformly spaced linear array (ULA) and attach a reference frame $\ubar{A} \triangleq (\udot{A},\uvec{A})$ to it, where $\udot{A}\in\mathbb{A}$ is its absolute position and $\uvec{A}\triangleq\big(\hat{x}_{A},\hat{y}_{A},\hat{z}_{A}\big)\in\mathbb{V}^3$ its absolute orientation (where $\hat{\cdot}$ denotes unit length). Assume that the array channels are digitally delayed to focus at a pixel with azimuth angle $\theta$ and range sample $n$, such that the delayed data from the $m$th channel can be expressed as $x_m[\theta,n-\Delta_m]\in\mathbb{C}$. We simplify notation by making the dependence on $\theta$ implicit from now on.

>>>>

Waves are made of endless tiny energy transactions, a dance of particles effecting each other in search of equilibrium. They are caused by disturbances in the forces of nature. The drop of a rock into calm water or rapid displacement of still air cause mechanical waves, while synchronized oscillations of electric and magnetic fields cause electromagnetic waves. Regardless of whether the waves are mechanical or electromagnetic, move perpendicular or parallel to the axis of propagation, what medium they propagate through or the means by which they do so, they are---perhaps surprisingly---all governed by a partial differential equation known as the \emph{wave equation}:
%
\begin{align}
\nabla^2 f 
 = \frac{1}{c}\frac{\delta^2 f}{\delta t^2},
\end{align}
%\frac{\delta^2 f}{\delta x^2} + \frac{\delta^2 f}{\delta y^2} + \frac{\delta^2 f}{\delta z^2}
%
where $\nabla^2$ is the Laplace operator, $f$ is field strength and $c$ is wave propagation speed. If $f$ represents the displacement of a vibrating string, the equation states that the force pulling the string towards equilibrium (right side) is proportional to the string's curvature (left side). In acoustics $f$ represents sound pressure, but since the mathematics are all the same we will simply think of $f$ as a signal: a unit-less scalar function with wave characteristics.

A few solutions to the wave equation is of interest, in particular that of a plane and spherical wave,
%
\begin{align}
f_\textrm{plane}(\vec{p},t)  &= A e^{j(\omega t - \vec{k} \cdot \vec{p})} \\
f_\textrm{spheric}(\vec{p},t) &= \frac{A}{\lVert \vec{p} \rVert} e^{j(\omega t - \lVert\vec{k}\rVert \lVert \vec{p} \rVert)},
\end{align}
% 
where $A$ is amplitude, $\omega$ is temporal wave frequency, $t$ is time, $\lVert\vec{k}\rVert=\frac{\omega}{c}=\frac{2\pi}{\lambda}$ is spatial wave frequency (wavenumber) and $\vec{p}$ is position of the source relative to the observer.

% One is that of plane wave and the other a  A signal source with the position $\udot{S}$ . We can represent the field observed at $\udot{A}$ as $f(\dvec{p}^A,t)$, where position is represented by $\dvec{p}^A=[x_A,y_A,z_A]^T\in\mathbb{R}^3$ and time by $t\in\mathbb{R}$,
% 
% An introduction to such notation can be found in \chapref{IV_methods}{IV}.
% 
% 
% reference frame $\ubar{A} \triangleq (\udot{A},\uvec{A})$ to it, where $\udot{A}\in\mathbb{A}$ is its absolute position and $\uvec{A}\triangleq\big(\hat{x}_{A},\hat{y}_{A},\hat{z}_{A}\big)\in\mathbb{V}^3$ its absolute orientation (where $\hat{\cdot}$ denotes unit length)

@@

Imagine we record a wavefield with $M$ sensors arranged in a uniformly spaced linear array (ULA) and attach a reference frame $\ubar{A} \triangleq (\udot{A},\uvec{A})$ to it, where $\udot{A}\in\mathbb{A}$ is its absolute position and $\uvec{A}\triangleq\big(\hat{x}_{A},\hat{y}_{A},\hat{z}_{A}\big)\in\mathbb{V}^3$ its absolute orientation (where $\hat{\cdot}$ denotes unit length). Assume that the array channels are digitally delayed to focus at a pixel with azimuth angle $\theta$ and range sample $n$, such that the delayed data from the $m$th channel can be expressed as $x_m[\theta,n-\Delta_m]\in\mathbb{C}$. By making the dependence on $\theta$ implicit and assuming the data to be pre-delayed to the pixel of interest, we can simply the notation to $x_m[\theta,n-\Delta_m] = x_m[n]$.

### 2019-08-22 ThesisJo/4_discussion.tex
2019-08-18
2019-08-21

Constrain attention to beamformers and simulators (not ATR?).

@@

>>>>

The contributions of this thesis are:

Paper I - Time-domain MVDR beamforming on GPUs in active sonar imaging:
\begin{enumerate}
\item Accelerates the MVDR beamformer through algorithmic reductions and implementation on GPU.
\item Demonstrates speeds exceeding 1\,Mpx/s on systems with no more than 32 channels, two orders of magnitude faster than a semi-optimized C-implementation on a CPU.
\item Offers uncommon insights into hardware limitations, design implications and computation efficiency.
\end{enumerate}

Paper II - Frequency-domain MVDR beamforming on GPUs in clinical ultrasound imaging:
\begin{enumerate}
\item Alters the implementation in Paper I to operate in a frequency domain approximated by its three most significant components.
\item Demonstrates speeds exceeding 1\,Mpx/s on systems with 64 and 96 channels.
\item Yields similar image quality to time-domain MVDR.
\end{enumerate}

Paper III - LCA beamforming compared to MVDR in active sonar imaging:
\begin{enumerate}
\item Contrasts LCA, that selects the best performing weights from a predefined set, to MVDR, that computes the weights from an estimated covariance matrix formed from the input data.
\item 
\item Yields similar image quality to MVDR.
\end{enumerate}

Paper IV - Real-time sonar image synthesis:
\begin{enumerate}
\item Contrasts LCA, that selects the best performing weights from a predefined set, to MVDR, that computes the weights from an estimated covariance matrix formed from the input data.
\item Extends a navigational notation system with topological ideas from robotics and homogeneous coordinates, promoting geometrical clarity and validity while aiding the designer in focusing on higher level concepts, such as manipulating rigid objects.
\item Yields similar image quality to MVDR.
\end{enumerate}

@@

\section{Onlook}

Machine cognition:
\begin{itemize}
\item Ethical dilemmas. Decision and actions have consequences. Who's legally responsible for a machine's doing? Itself?
\item Safety. Machine flexibility allow more possible outcomes, and a wider range of security issues to consider.
\end{itemize}

R dim > 10000
- Inverting such matrices better to do with gradient decent

Deep learning. Simulator used to trained deep neural networks. Lots of data needed.

\section{}
